---
title: "A quick guide to scAMACE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A quick guide to scAMACE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Datasets
### Application 1: K562-GM12878 dataset
`Feb7_2021_3Types_Data_rna_mean_1000_ratio.RData`: Application 1: K562-GM12878 dataset, including scCAS data, scRNA-Seq data and sc-methylation data.

`Feb7_2021_3Types_Data_rna_mean_1000_ratio_mcmc_ini.RData`: Initialization for Application 1: K562-GM12878 dataset, including initialization data for scCAS data, scRNA-Seq data and sc-methylation data.


### Application 2: Mouse neocortex dataset
`Feb9_2021_new_three_data_5subcell_Astro_CPN_L6_mPv_oligo_1000gene_ratio.RData`: Application 2: Mouse neocortex dataset, including scCAS data, scRNA-Seq data and sc-methylation data.

`App2_Feb9_2021_new_three_data_5subcell_Astro_CPN_L6_mPv_oligo_complicated_rna_vtilde_all_link_real_3data_ini_square.RData`: Initialization for Application 2: Mouse neocortex dataset, including initialization data for scCAS data, scRNA-Seq data and sc-methylation data.


## 2. Example
## Remarks: We demostrate usage of scAMACE through Application 2.
### 2.1 Load data and prepare for EM algorithm
```{r}
library(scAMACE)

### load the data
data('Feb9_2021_new_three_data_5subcell_Astro_CPN_L6_mPv_oligo_1000gene_ratio')
data <- Feb9_2021_new_three_data_5subcell_Astro_CPN_L6_mPv_oligo_1000gene_ratio


f10ratio_acc <- data$f10ratio_acc
f1 <- data$f1
f0 <- data$f0
atac_cell_lb <- data$atac_cell

f10ratio_exp <- data$f10ratio_exp
g1 <- data$g1
g0 <- data$g0
rna_cell_lb <- data$rna_cell

f10ratio_met <- data$f10ratio_met
met_data <- data$met_data
met_cell_lb <- data$met_cell


### load the initialization
data("App2_Feb9_2021_new_three_data_5subcell_Astro_CPN_L6_mPv_oligo_complicated_rna_vtilde_all_link_real_3data_ini_square")
ini <- Feb9_2021_complicated_rna_vtilde_all_link_real_3data_ini


###----------------------------------------------------------------------###
k0 <- 5
i0 <- nrow(f10ratio_acc)
l0 <- nrow(f10ratio_exp)
d0 <- nrow(f10ratio_met)
p0 <- ncol(f10ratio_acc)


niter <- 200


rij <- f1/(f1+f0)
replacef0 <- 1-rij

rlg <- g1/(g1+g0)
replaceg0 <- 1-rlg


ratio <- met_data/(1-met_data)
c <- ratio[which(ratio!=0)]

sf <- quantile(c,probs = 0.5,na.rm=T) # median
sf
rdm <- ratio/sf
rdm <- rdm^2 # power
rdm[is.na(rdm)] <- mean(rdm,na.rm = TRUE)
rdm[1:5,1:5]
replaceh0 <- matrix(1,nrow=d0,ncol=p0)



alpha1 <- 2
beta1 <- 2


alpha_qi <- 1
beta_qi <- 1


### beta regression
###----------------------------------------------------------------------###
phi_1 <- 5.043486
eta <- -2.762517    
gamma <-  4.106395
tau <- -2.642951


phi_2 <- 0.8038949
delta <- -0.7348618 
theta <- 2.0083089 




### initialization
###----------------------------------------------------------------------###
phi_atac <- rep(1/k0,k0)
phi_rna <- rep(1/k0,k0)
phi_met <- rep(1/k0,k0)

w <- ini$omega_exp
w_acc <- ini$omega_acc
w_acc <- w
w_exp <- ini$omega_exp
w_exp <- w
w_met <- ini$omega_met
w_met <- w

pi_rna <- ini$pi_exp
pi_met <- ini$pi_met
qi <- ini$qi_acc



```


### 2.2 run the EM algorithm
```{r}
print("start EM")
start_time <- Sys.time()
temp<-cal_M_step(phi_atac=phi_atac,w_acc=w_acc,qi=qi,
                  phi_rna=phi_rna,w_exp=w_exp,pi_rna=pi_rna,
                  phi_met=phi_met,w_met=w_met,pi_met=pi_met)

for(i in c(1:niter)){
  temp<-cal_M_step(phi_atac=temp$phi_atac,w_acc=temp$w_acc,qi=temp$qi,
                    phi_rna=temp$phi_rna,w_exp=temp$w_exp,pi_rna=temp$pi_rna,
                    phi_met=temp$phi_met,w_met=temp$w_met,pi_met=temp$pi_met)
  print(i)
  
}

pst <- cal_post(phi_atac=temp$phi_atac,w_acc=temp$w_acc,qi=temp$qi,
                phi_rna=temp$phi_rna,w_exp=temp$w_exp,pi_rna=temp$pi_rna,
                phi_met=temp$phi_met,w_met=temp$w_met,pi_met=temp$pi_met)
pst

end_time <- Sys.time()
end_time - start_time

print("end EM")

```


### 2.3 Summary clustering result (get the cluster assignments)
```{r}
prob_z_atac<-cal_E_acc(rij,replacef0,temp$phi_atac,temp$w_acc,temp$qi)$prob_z_atac
prob_z_rna<-cal_E_rna(rlg,replaceg0,temp$phi_rna,temp$w_exp,temp$pi_rna)$prob_z_rna
prob_z_met<-cal_E_rna(rdm,replaceh0,temp$phi_met,temp$w_met,temp$pi_met)$prob_z_rna

# amp, rmp and mmp are the cluster assignments offered by scAMACE
# scCAS data
amp<-rep(0,i0)
for(i in c(1:i0)){
  amp[i]=which.max(prob_z_atac[i,])
}
amp[1:10]

# scRNA-Seq data
rmp<-rep(0,l0)
for(l in c(1:l0)){
  rmp[l]=which.max(prob_z_rna[l,])
}
rmp[1:10]

# sc-methylation data
mmp<-rep(0,d0)
for(d in c(1:d0)){
  mmp[d]=which.max(prob_z_met[d,])
}
mmp[1:10]


### contingency table
# table(atac_cell_lb,amp)

# table(rna_cell_lb,rmp)

# table(met_cell_lb,mmp)

```



